---
title: "DPAS光老化量表的重现性研究"
author: "Lian Rui"
date: "2026-09-24"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  eval = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	include = TRUE,
  results = "markup",
  fig.show = "asis"
)
```

## 1. Setup

### Packages

```{r packages, include=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(tidyr)
library(psych) # ICC
library(ggplot2)
library(colorspace)
library(showtext)
showtext_auto()
library(GWalkR) # Tabuleu like, TODO: furrther explore
library(DataExplorer) 
library(gtsummary) # Table 1
library(kableExtra)
library(lmerTest)
```

### Global options

### Helper functions

```{r}
utils.path <- "~/DPAS Reproducibility/utils"
```


```{r}
source(file.path(utils.path, "calculate_total_icc.R"))

source(file.path(utils.path, "calculate_metric_icc.R"))
```


## 2. Data import

### Path

```{r path}
path <- "~/DPAS Reproducibility/data"
```

### raw_dpas

```{r raw_dpas import}
raw_dpas <- read_excel(
  file.path(path, "DPAS48.xlsx")
)
```

### raw_pt

```{r raw_pt import}
raw_pt <- read_excel(
  file.path(path, "DPAS48_pt_info.xls")
)
```

## 3. Data cleaning

### clean_dpas

```{r clean DPAS col names}
clean_dpas <- raw_dpas |>
  rename(
    Patient_id = 患者,
    Rater = 监察人
  ) |>
  rename_with(
    ~ paste0("Forehead", sub("^额", "", .x)),
    matches("^额\\d+$")
  ) |>
  rename_with(
    ~ paste0("Right_cheek", sub("^右", "", .x)),
    matches("^右\\d+$")
  ) |>
  rename_with(
    ~ paste0("Left_cheek", sub("^左", "", .x)),
    matches("^左\\d+$")
  ) |>
  rename_with(
    ~ paste0("Chin", sub("^下", "", .x)),
    matches("^下\\d+$")
  )
```

### clean_dpas_long

```{r clean DPAS long format}
clean_dpas_long <- clean_dpas |>
  pivot_longer(
    cols = 3:ncol(clean_dpas),
    names_to = "Position_metric",
    values_to = "Score"
  ) |>
  separate(
    Position_metric,
    into = c("Position", "Metric"),
    sep = "(?<=\\D)(?=\\d)",
    convert = TRUE
  )
```

### clean_pt

```{r rename revalue raw_pt}
clean_pt <- raw_pt |>
  rename(
    Patient_id = 姓名,
    Sex = 性别,
    Age = 年龄
  ) |>
  mutate(
    Sex = recode(
      Sex,
      "男" = "Male",
      "女" = "Female"
    )
  )
```

### Create canonical(single source of truth) dataset dat_long

```{r Merge DPAS and pt_info}
dat_long <- clean_pt |>
  left_join(
    clean_dpas_long,
    by = "Patient_id"
  )
```

### Create dat_model for logistic mixed model

```{r dat_model}
dat_model <- dat_long |>
  mutate(
    Age_c = Age - mean(Age, na.rm = TRUE),
    Metric = factor(Metric),
    Position = factor(Position),
    Sex = factor(Sex)
  )
```


## 4. Data quality check

### Missing values

```{r missing value check}
sapply(dat_long, function(x) sum(is.na(x)))
```
### internal structure of dat_long

```{r dat_long structure}
str(dat_long)
```

### Change data type of each column

```{r change column class}
dat_long <- dat_long %>%
  mutate(
    Patient_id = as.character(Patient_id),
    Sex        = factor(Sex),
    Age        = as.numeric(Age), # already numeric → no-op, kept for clarity
    Rater      = factor(Rater),
    Position   = factor(Position),
    Metric     = factor(Metric),
    Score      = as.numeric(Score) # already numeric → no-op, kept for clarity
  )
```


### Data summary

```{r}
summary(dat_long)
```

## 5. Descriptive analysis

In `dat_long`, the columns are: 

```{r colnames dat long}
colnames(dat_long)
```
### Overall framework of descriptive analysis

We can define `Sex`, `Age`, `Rater`, `Position`, `Metric` as predictors and `Score` as outcome. and the description analysis in nature is combination of predictors and the distribution, mean, count etc of outcome. And such predictor combination has some hierarchy, e.g., single variable, bivariates and so on. Create a framework of description analysis

A general, combinatary structure as below: 

```text
Level 0: Outcome alone
        │
        ├── Overall distribution of Score
        │
Level 1: Single predictor
        │
        ├── Sex
        ├── Age
        ├── Rater
        ├── Position
        └── Metric
        │
Level 2: Two-dimensional descriptions
        │
        ├── Predictor × Score
        ├── Predictor × Predictor
        └── clinically/design-relevant combinations
        │
Level 3: Three-dimensional descriptions
        │
        ├── Sex × Position × Score
        ├── Sex × Metric × Score
        ├── Rater × Position × Score
        └── Rater × Metric × Score
        │
Level 4: Full / high-dimensional structure
        │
        └── Sex × Age × Rater × Position × Metric × Score
```

But this hierarchy makes the description analysis cumbersome and many of them are not scientific valid. 

### The scientific assumptions and questions behind the data

1. Sex related question: What are the difference of skin photoaging between male and female? 

1.1 Total score difference between male and female

1.2 The total score by facial area (chin, forehead, right-left cheek)

1.3 The overall lesion type difference

1.4 The lesion type by facial area

2. Age related question

2.1 The correlation of score with age and sex

2.2 The score of each facial area correlated with age and sex

2.3 the possible lesion type difference among young and old group and sex

3. Rater related question

3.1 The four scores of each patient by four raters

3.2 The overall distribution of total score by 4 raters

3.3 The distribution of facial area score by 4 raters

3.4 The lesion by four raters

3.5 The lesion by four raters by facial area

### Table 1

```{r Table 1}
table1_df <- dat_long |>
  dplyr::select(Patient_id, Sex, Age) |>
  dplyr::distinct(Patient_id, .keep_all = TRUE) |>
  gtsummary::tbl_summary(
    by = Sex,
    include = Age,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    missing = "ifany"
  ) |>
  gtsummary::add_p() |>
  tibble::as_tibble()

table1_df
```


### Distributions of score by rater sex

```{r score by rater sex, warning=FALSE}
total_score_by_rater_sex <- dat_long |>
  group_by(Patient_id, Sex, Age, Rater) |>
  summarise(total = sum(Score, na.rm = TRUE)) |>
  ungroup()

# Visualize the distribution of total score
#| fig-width: 5
#| fig-height: 3
#| out-width: 70%
g<- ggplot(total_score_by_rater_sex, aes(x = total)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "grey") +
  labs(title = "Distribution of Total Score by rater and sex", x = "Total Score", y = "Count") +
  theme_minimal() + 
  facet_grid(Rater ~ Sex)

g
```

### Mean score by age, sex

```{r average score}
avg_score <- total_score_by_rater_sex |>
  group_by(Patient_id, Age, Sex) |>
  summarise(avg = mean(total, na.rm = TRUE))
```

```{r avg score boxplot by sex}
g <- ggplot(avg_score, aes(x = Sex, y= avg, color = Sex)) + 
  geom_boxplot(show.legend = FALSE) + 
  geom_point(show.legend = FALSE) + 
  labs(title = "患者DPAS平均分的分布情况", 
       subtitle = "患者分数由四个评估人的分数平均得到",
       x = "性别",
       y = "DPAS 平均分") + 
  theme_minimal()

g
```

```{r compare mean by sex}
dpas_mean <- avg_score |>
  group_by(Sex) |>
  summarise(sex_mean = mean(avg, na.rm = TRUE), 
            sd = sd(avg, na.rm = TRUE))

g <- ggplot(dpas_mean, aes(x = Sex, y = sex_mean)) + 
  geom_col(aes(fill = Sex),
           show.legend = FALSE) + 
  geom_errorbar(aes(ymin = sex_mean - sd, ymax = sex_mean + sd), 
                width = 0.1) + 
  labs(title = "患者DPAS评分均值±SD",
       y = "患者DPAS评分均值",
       x = "性别") + 
  theme_minimal()

g
```


```{r correlation with age}

g <- ggplot(avg_score, aes(x = Age, y= avg, colour = Sex)) + 
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Mean DPAS Score by Age and Sex", x = "Age", y = "Mean DPAS Score") +
  theme_minimal()

g
```

### 男女在11项评分中的表现和差别

```{r metric by sex}
avg_score_sex_metric <- dat_long |>
  group_by(Sex, Position, Metric) |>
  summarise(avg_score = mean(Score)) |>
  ungroup()


g <- ggplot(avg_score_sex_metric, 
            aes(x = Metric, y = avg_score, fill = Sex)) + 
  geom_col(position = position_dodge()) + 
  facet_wrap(Position~., nrow = 2) + 
  theme_minimal()

g
```

## 6. Analysis dataset

All analysis dataset derived from canonical `dat_long`

```text
dat_long --> |group by id, rater, sex, age, sum| --> dat_total_rater

  dat_total_rater --> |group y id, sex, age, mean| --> dat_mean_patient

```

### dat_total_rater

```{r total score}
dat_total_rater <- dat_long |>
  group_by(Patient_id, Rater, Sex, Age) |>
  summarise(
    Total = sum(Score, na.rm = TRUE),
    .groups = "drop"
  )
```

### dat_metric_specific_rater

```{r metric specific score}
dat_metric_specific_rater <- dat_long |>
  group_by(Patient_id, Rater, Sex, Age, Metric) |>
  summarise(
    Total = sum(Score, na.rm = TRUE),
    .groups = "drop"
  )
```


### dat_mean_patient

```{r mean of pt score}
dat_mean_patient <- dat_total_rater |>
  group_by(Patient_id, Sex, Age) |>
  summarise(
    mean_total = mean(Total, na.rm = TRUE),
    .groups = "drop"
  )
```


## 7. Intraclass correlation: reliability of raters

### Total score ICC

```{r total icc}
icc_total <- calculate_total_icc(dat_total_rater)

icc_total$results
```

### Metric specific ICC

```{r metric icc}
icc_metric <- calculate_metric_icc(dat_metric_specific_rater)

icc_metric
```

## 8. Linear Mixed effect model

```{r glmer fit, echo=FALSE}
fit <- glmer(
  Score ~ Metric +  Position +  Sex + Age_c + Metric * Sex + 
    (1 | Rater) + (1 | Patient_id),
  data = dat_model,
  family = binomial
)

# 3. Check convergence
isSingular(fit)
summary(fit)
```

