---
title: "Untitled"
author: "Lian Rui"
date: "2026-07-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Workflow

```text
                    Raw Excel
                        │
                        ▼
                 Read Excel file
                        │
        ┌───────────────┴───────────────┐
        ▼                               ▼
 Total DPAS score                 44 Individual Items
        │                               │
        ▼                               ▼
 Overall ICC                  ICC of each item
        │
        ├──────────────┐
        ▼              ▼
Regional ICC      Metric ICC
        │              │
        └──────┬───────┘
               ▼
        Summary Table
               │
               ▼
          Heatmap
               |
               ▼
    Lowest agreement items
```

## Packages

```{r}
# packages
library(readxl)
library(dplyr)
library(tidyr)
library(psych) # ICC
library(ggplot2)
```

## Step 1: Getting The Data

```{r}
# read Excel
dat <- read_excel("data/Book1.xlsx")
```

## Step 2: Define four Facial Regions:Forehead, Left Cheek, Right Cheek, Chin

```{r}
# columns for each region
regions <- list(
  Forehead = 3:13,
  LeftCheek = 14:24,
  RightCheek = 25:35,
  Chin = 36:46
)
```

## Step 3: Define 11 metrics

```{r}
# columns for each metric
metric_map <- list(
  Metric1 = c(3, 14, 25, 36),
  Metric2 = c(4, 15, 26, 37),
  Metric3 = c(5, 16, 27, 38),
  Metric4 = c(6, 17, 28, 39),
  Metric5 = c(7, 18, 29, 40),
  Metric6 = c(8, 19, 30, 41),
  Metric7 = c(9, 20, 31, 42),
  Metric8 = c(10, 21, 32, 43),
  Metric9 = c(11, 22, 33, 44),
  Metric10 = c(12, 23, 34, 45),
  Metric11 = c(13, 24, 35, 46)
)
```

## Step 4: Explore The Distribution of Total Score

```{r}
# Calculate total score of each subject
dat$total <- rowSums(dat[, 3:46])
```

### Visualize the distribution of total score

```{r}
# Visualize the distribution of total score
ggplot(dat, aes(x = total)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "grey") +
  labs(title = "Distribution of Total Score", x = "Total Score", y = "Count") +
  theme_minimal()
```
### Visualize the distribution of total score by rater

```{r}
# The distribution of total score by rater
overall <- dat %>%
  select(患者, 监察人, total) %>%
  pivot_wider(
    names_from = 监察人,
    values_from = total
  )
```

```{r}
# Visualize the distribution of total score by rater
overall_long <- overall %>%
  pivot_longer(
    cols = -患者,
    names_to = "Rater",
    values_to = "TotalScore"
  )

ggplot(overall_long, aes(x = TotalScore, fill = Rater)) +
  geom_histogram(binwidth = 1, position = "dodge", color = "white") +
  labs(title = "Distribution of Total Score by Rater", x = "Total Score", y = "Count") +
  facet_grid(Rater ~ .) +
  theme_minimal()
```

### Visualize the overall score of each subject

```{r}
ggplot(overall_long, aes(x = factor(患者), y = TotalScore, 
                         color = Rater, shape = Rater)) +
  geom_point(size = 3) +
  labs(title = "Overall Score of Each Subject by Rater", x = "Subject", y = "Total Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Step 5: The ICC of overall score

```{r}
# Calculate ICC for overall score
overall_icc <- ICC(overall[, -1])

overall_icc$results
```

### TODO: Explain the ICC results

## Step 6: Calculate ICC for each region

### Question: The ICC of each region, any differences?

```{r}
region_results <- data.frame()

# TODOs: Annotation of the codes below.

for (r in names(regions)) {
  
  dat$score <- rowSums(dat[, regions[[r]]])

  tmp <- dat %>%
    select(患者, 监察人, score) %>%
    pivot_wider(
      names_from = 监察人,
      values_from = score
    )

  icc <- ICC(tmp[, -1])

  region_results <- rbind(
    region_results,
    data.frame(
      Region = r,
      ICC = icc$results["Single_random_raters", "ICC"]
    )
  )
}

region_results
```

### TOTO: Explain the ICC results by region

### Question: any differences in the scores by rater and by region?

```{r}
# Create a long format data frame for plotting
long_dat <- dat %>%
  pivot_longer(
    cols = 3:46,
    names_to = "Item",
    values_to = "Score"
  )

# Summarize the scores by rater and region
summary_scores <- long_dat %>%
  mutate(Region = case_when(
    Item %in% names(dat)[regions$Forehead] ~ "Forehead",
    Item %in% names(dat)[regions$LeftCheek] ~ "LeftCheek",
    Item %in% names(dat)[regions$RightCheek] ~ "RightCheek",
    Item %in% names(dat)[regions$Chin] ~ "Chin"
  )) %>%
  group_by(监察人, Item) %>%
  summarise(Average_Score = mean(Score, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
# Visualize the average score by rater and region
ggplot(summary_scores, aes(x = 监察人, y = Average_Score, fill = Item)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Average Scores by Rater and Region",
    x = "监察人", y = "Average Score"
  ) +
  facet_grid(Item ~ .) +
  theme_minimal()
```

## Step 5 Calculate ICC for each metric

### Question: The ICC of each metric by rater

```{r}
metric_long <- metric_results %>%
  pivot_longer(
    -Metric,
    names_to = "Rater",
    values_to = "MeanScore"
  ) %>%
  mutate(
    Metric = factor(
      Metric,
      levels = paste0("Metric", 1:11)
    )
  )
```


```{r}
ggplot(
  metric_long,
  aes(Rater, Metric, fill = MeanScore)
) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.2f", MeanScore)), size = 4) +
  scale_y_discrete(limits = rev) + # Metric1 at top
  scale_fill_gradient2(
    low = "steelblue",
    mid = "white",
    high = "firebrick",
    midpoint = mean(metric_long$MeanScore)
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Mean\nScore"
  ) +
  theme_minimal(base_size = 14)
```
