---
title: "Untitled"
author: "Lian Rui"
date: "2026-07-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Version log

Add mixed-effect model 

Revise the code

## Workflow

### TODO: this flow diagram need to be updated

```text
                    Raw Excel
                        │
                        ▼
                 Read Excel file
                        │
        ┌───────────────┴───────────────┐
        ▼                               ▼
 Total DPAS score                 44 Individual Items
        │                               │
        ▼                               ▼
 Overall ICC                  ICC of each item
        │
        ├──────────────┐
        ▼              ▼
Regional ICC      Metric ICC
        │              │
        └──────┬───────┘
               ▼
        Summary Table
               │
               ▼
          Heatmap
               |
               ▼
    Lowest agreement items
```

## Packages

```{r}
# packages
library(tidyverse)
library(readxl)
library(dplyr)
library(tidyr)
library(psych) # ICC
library(ggplot2)
library(colorspace)
```

## Step 1: Getting The Data

```{r}
# read Excel
dat <- read_excel("data/Book1.xlsx")
```

### Long format data

```{r}
dat_long <- dat %>%
  pivot_longer(
    cols = 3:46,
    names_to = "Item",
    values_to = "Score"
  ) %>%
  mutate(
    Site = case_when(
      str_detect(Item, "^额") ~ "Forehead",
      str_detect(Item, "^左") ~ "LeftCheek",
      str_detect(Item, "^右") ~ "RightCheek",
      str_detect(Item, "^下") ~ "Chin"
    ),

    Metric = str_extract(Item, "\\d+$"),

    Patient = factor(患者),
    Rater = factor(监察人),
    Site = factor(Site),
    Metric = factor(Metric)
  ) %>%
  select(Patient, Rater, Site, Metric, Score)
```

## The total score of each patient

```{r}
# Calculate total score of each patient from dat_long
total_score <- dat_long %>%
  group_by(Patient, Rater) %>%
  summarise(TotalScore = sum(Score, na.rm = TRUE)) %>%
  ungroup()
```

### Visualize the distribution of total score: Histogram

```{r}
# Visualize the distribution of total score
ggplot(total_score, aes(x = TotalScore)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "grey") +
  labs(title = "Distribution of Total Score", x = "Total Score", y = "Count") +
  theme_minimal() + 
  facet_grid(Rater ~ .)
```

### Visualize the distribution of total score: density plot

```{r}
# Visualize the distribution of total score as a density plot 
ggplot(total_score, aes(x = TotalScore)) +
  geom_density(fill = "steelblue", color = "grey") +
  labs(title = "Distribution of Total Score", x = "Total Score", y = "Count") +
  geom_vline(xintercept = c(7, 10), color = "orange", linetype = "dashed", size = 1) +
  theme_minimal() + 
  facet_grid(Rater ~ .)
```

患者的评分集中在7-10分之间，说明大部分患者的评分在这个范围内。

三个评分人的评分分布大致相似，说明评分人之间的评分一致性较高。

## The ICC of overall score

Calculate the intraclass correlation(ICC) of three raters for the overall score of each patient.

```{r}
# convert the total_score data frame to wide format for ICC calculation
icc_data <- total_score %>%
  pivot_wider(
    names_from = Rater,
    values_from = TotalScore,
    names_prefix = "Rater"
  )

# Calculate the icc using the psych package
icc_result <- ICC(icc_data[, -1]) # Exclude the Patient column 

View(icc_result$results)
```

### TODO: Explain the ICC results

## 对11项指标，三个评分人的评分是否存在主观差异

### Visualization

```{r}
metric_long <- dat_long %>%
  group_by(Metric, Rater) %>%
  summarise(Average_Score = mean(Score, na.rm = TRUE)) %>%
  ungroup()

metric_long$Metric <- factor(metric_long$Metric, levels = (1:11))
```

```{r}
# Visualize the average score of each metric by rater
# The metric ordered from 1 to 1-10
ggplot(metric_long, aes(x = Rater, 
                        y = fct_rev(Metric),
                        fill = Average_Score)) +
  geom_tile(color = "grey") +
  geom_text(aes(label = round(Average_Score, 2)), 
            color = "steelblue", 
            size = 4) + 
  labs(title = "Average Score of Each Metric by Rater", 
       x = "Metric", y = "Average Score") +
  scale_fill_continuous_diverging(palette = "Cyan-Magenta") +
  theme_minimal()

```

```{r}
library(dplyr)
library(tidyr)
library(psych)

icc_results <- data.frame()

for (m in sort(unique(dat_long$Metric))) {

  tmp <- dat_long %>%
    filter(Metric == m) %>%
    group_by(Patient, Rater) %>%
    summarise(
      Score = sum(Score),    # sum across the 4 facial sites
      .groups = "drop"
    ) %>%
    pivot_wider(
      names_from = Rater,
      values_from = Score
    )

  icc <- ICC(tmp[, -1])

  icc_results <- bind_rows(
    icc_results,
    data.frame(
      Metric = m,
      ICC2_1 = icc$results["Single_random_raters", "ICC"],
      Lower = icc$results["Single_random_raters", "lower bound"],
      Upper = icc$results["Single_random_raters", "upper bound"],
      p = icc$results["Single_random_raters", "p"]
    )
  )
}

icc_results
```



#####################

## Calculate ICC for 11 metrics 

```{r}
# 
icc_metric <- dat_long %>%
  group_by(Metric) %>%
  summarise(
    icc = ICC(select(., Score), model = "twoway", type = "agreement")$results["Single_random_raters", "ICC"]
  ) %>%
  ungroup()
```


### Question: The ICC of each region, any differences?

```{r}
region_results <- data.frame()

# TODOs: Annotation of the codes below.

for (r in names(regions)) {
  
  dat$score <- rowSums(dat[, regions[[r]]])

  tmp <- dat %>%
    select(患者, 监察人, score) %>%
    pivot_wider(
      names_from = 监察人,
      values_from = score
    )

  icc <- ICC(tmp[, -1])

  region_results <- rbind(
    region_results,
    data.frame(
      Region = r,
      ICC = icc$results["Single_random_raters", "ICC"]
    )
  )
}

region_results
```

### TOTO: Explain the ICC results by region

### Question: any differences in the scores by rater and by region?

```{r}
# Create a long format data frame for plotting
long_dat <- dat %>%
  pivot_longer(
    cols = 3:46,
    names_to = "Item",
    values_to = "Score"
  )

# Summarize the scores by rater and region
summary_scores <- long_dat %>%
  mutate(Region = case_when(
    Item %in% names(dat)[regions$Forehead] ~ "Forehead",
    Item %in% names(dat)[regions$LeftCheek] ~ "LeftCheek",
    Item %in% names(dat)[regions$RightCheek] ~ "RightCheek",
    Item %in% names(dat)[regions$Chin] ~ "Chin"
  )) %>%
  group_by(监察人, Item) %>%
  summarise(Average_Score = mean(Score, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
# Visualize the average score by rater and region
ggplot(summary_scores, aes(x = 监察人, y = Average_Score, fill = Item)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Average Scores by Rater and Region",
    x = "监察人", y = "Average Score"
  ) +
  facet_grid(Item ~ .) +
  theme_minimal()
```

## Step 5 Calculate ICC for each metric

### Question: The ICC of each metric by rater

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

## Calculate average score of each metric for each rater
metric_results <- lapply(names(metric_map), function(metric){

  dat %>%
    mutate(
      metric_score = rowMeans(
        select(., all_of(metric_map[[metric]])),
        na.rm = TRUE
      )
    ) %>%
    group_by(监察人) %>%
    summarise(
      MeanScore = mean(metric_score),
      .groups = "drop"
    ) %>%
    mutate(Metric = metric)

}) %>%
  bind_rows()
```


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

metric_long <- metric_results %>%
  pivot_longer(
    cols = -Metric,
    names_to = "Rater",
    values_to = "MeanScore"
  ) %>%
  mutate(
    Metric = factor(
      Metric,
      levels = paste0("Metric", 1:11)
    )
  )

ggplot(metric_long,
       aes(x = Rater,
           y = Metric,
           fill = MeanScore)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.2f", MeanScore)), size = 4) +
  scale_y_discrete(limits = rev(levels(metric_long$Metric))) +
  scale_fill_gradient2(
    low = "steelblue",
    mid = "white",
    high = "firebrick",
    midpoint = mean(metric_long$MeanScore)
  ) +
  labs(
    x = "Rater",
    y = "Metric",
    fill = "Mean Score"
  ) +
  theme_minimal(base_size = 14)
```


```{r}
ggplot(
  metric_long,
  aes(Rater, Metric, fill = MeanScore)
) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = sprintf("%.2f", MeanScore)), size = 4) +
  scale_y_discrete(limits = rev) + # Metric1 at top
  scale_fill_gradient2(
    low = "steelblue",
    mid = "white",
    high = "firebrick",
    midpoint = mean(metric_long$MeanScore)
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Mean\nScore"
  ) +
  theme_minimal(base_size = 14)
```


## Mixed effect model

```{r}
library(tidyverse)

dat <- readxl::read_excel("data/Book1.xlsx")

long <- dat %>%
  pivot_longer(
    cols = 3:46,
    names_to = "Item",
    values_to = "Score"
  ) %>%
  mutate(
    Site = case_when(
      str_detect(Item, "^额") ~ "Forehead",
      str_detect(Item, "^左") ~ "LeftCheek",
      str_detect(Item, "^右") ~ "RightCheek",
      str_detect(Item, "^下") ~ "Chin"
    ),

    Metric = str_extract(Item, "\\d+$"),

    Patient = factor(患者),
    Rater = factor(监察人),
    Site = factor(Site),
    Metric = factor(Metric)
  ) %>%
  select(Patient, Rater, Site, Metric, Score)
```

```{r}
library(lme4)

fit <- glmer(
  Score ~ Site + Metric * Rater +
    (1 | Patient), 
  family = binomial,
  data = long
)

summary(fit)
anova(fit)
```

