---
title: "Untitled"
author: "Lian Rui"
date: "2026-07-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
88888888889 `

## Workflow


```text
                    Raw Excel
                        │
                        ▼
                 Read Excel file
                        │
        ┌───────────────┴───────────────┐
        ▼                               ▼
 Total DPAS score                 11 Individual Items
        │                               │
        ▼                               ▼
 Overall ICC Between raters       ICC of each item
        |_______________________________|
                        |
                Mixed effect model
    
```

## Packages

```{r}
# packages
library(tidyverse)
library(readxl)
library(dplyr)
library(tidyr)
library(psych) # ICC
library(ggplot2)
library(colorspace)
```

## Step 1: Getting The Data

```{r}
# read Excel
dat <- read_excel("data/副本Book1.xlsx")
```

### Transform To Long Format Data

```{r}
dat_long <- dat %>%
  pivot_longer(
    # Column 3 to 46 are the 44 items
    cols = 3:46,
    
    # The names of the new columns
    names_to = "Item",
    
    # The values of the new columns
    values_to = "Score"
  ) %>%
  mutate(
    Site = case_when(
      # Use regex to match the first character of the Item column
      
      # If the first character is "额", then the Site is "Forehead"
      str_detect(Item, "^额") ~ "Forehead",
      
      # If the first character is "左", then the Site is "LeftCheek"
      str_detect(Item, "^左") ~ "LeftCheek",
      
      # If the first character is "右", then the Site is "RightCheek"
      str_detect(Item, "^右") ~ "RightCheek",
      
      # If the first character is "下", then the Site is "Chin"
      str_detect(Item, "^下") ~ "Chin"
    ),
    # Extract the last digit of the Item column as the Metric
    Metric = str_extract(Item, "\\d+$"),
    
    # Convert the Patient, Rater, Site, and Metric columns to factors
    Patient = factor(患者),
    Rater = factor(监察人),
    Site = factor(Site),
    Metric = factor(Metric)
  ) %>%
  select(Patient, Rater, Site, Metric, Score)
```

## The total score of each patient

```{r}
# Calculate total score of each patient from dat_long
total_score <- dat_long %>%
  group_by(Patient, Rater) %>%
  summarise(TotalScore = sum(Score, na.rm = TRUE)) %>%
  ungroup()
```

### Visualize the distribution of total score: Histogram

```{r}
# Visualize the distribution of total score
ggplot(total_score, aes(x = TotalScore)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "grey") +
  labs(title = "Distribution of Total Score", x = "Total Score", y = "Count") +
  theme_minimal() + 
  facet_grid(Rater ~ .)
```


患者的评分集中在7-10分之间，说明大部分患者的评分在这个范围内。

三个评分人的评分分布大致相似，说明评分人之间的评分一致性较高。

## The ICC of overall score

Calculate the intraclass correlation(ICC) of three raters for the overall score of each patient.

```{r}
# convert the total_score data frame to wide format for ICC calculation
icc_data <- total_score %>%
  pivot_wider(
    # The names of the new columns
    names_from = Rater,
    
    # The values of the new columns
    values_from = TotalScore,
    
    # The prefix for the new columns
    names_prefix = "Rater"
  )

# Calculate the icc using the psych package
icc_result <- ICC(icc_data[, -1]) # Exclude the Patient column 
```

```{r}
icc_result$results
```


## 对11项指标，三个评分人的评分是否存在主观差异

```{r}
# Calculate the ICC for each metric
# A data frame to store the ICC results
icc_results <- data.frame()

# Loop through each unique metric
for (m in seq_along(unique(dat_long$Metric))) {
  
  # tmp is a temporary data frame that contains the scores for the current metric
  tmp <- dat_long %>%
    # metric filtered as m in the loop
    filter(Metric == m) %>%
    group_by(Patient, Rater) %>%
    summarise(
      # for each metric, sum the scores across the 4 facial sites
      Score = sum(Score),    # sum across the 4 facial sites
      # remove the grouping structure after summarising
      .groups = "drop"
    ) %>%
    # Transform to wide format for ICC calculation
    pivot_wider(
      names_from = Rater,
      values_from = Score
    )

  icc <- ICC(tmp[, -1])
  # Store the ICC results in the icc_results data frame
  icc_results <- bind_rows(
    icc_results,
    data.frame(
      Metric = m,
      # Extract the ICC value for single random raters from the ICC results
      ICC2_1 = icc$results["Single_random_raters", "ICC"],
      
      # Extract the upper and lower bound of the ICC value 
      # for single random raters from the ICC results
      Lower = icc$results["Single_random_raters", "lower bound"],
      Upper = icc$results["Single_random_raters", "upper bound"],
      
      # p value
      p = icc$results["Single_random_raters", "p"]
    )
  )
}

icc_results
```


## Mixed effect model


```{r}
library(lme4)

fit <- glmer(
  Score ~ Site + Metric +
    (1 | Rater) + (1 | Patient), 
  family = binomial,
  data = dat_long
)

summary(fit)
anova(fit)
```

