---
title: "皮肤镜下光老化量表的重现性研究"
author: "xxx，xxx，xxx"
date: "2026-09-24"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
bibliography: references.bib
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, # hide R source code from the final document, but run it
  eval = TRUE, # Actually execute the code
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	include = TRUE,
  results = "markup",
  fig.show = "asis" # All figures placed where they are produced, 
  ## hold: all figures are held and placed at the end of the chunk
  ## Animate: combine multiple plots into an animation!
)

options(digits = 3)
```


```{r packages}
library(tidyverse)
library(readxl)
library(dplyr)
library(tidyr)
library(psych) # ICC
library(ggplot2)
library(colorspace)
library(showtext)
showtext_auto()
library(GWalkR) # Tabuleu like, TODO: further explore
library(DataExplorer) 
library(gtsummary) # Table 1
library(kableExtra)
library(lmerTest)
library(emmeans)
library(forcats)
library(scales)
library(flextable)
library(officer)
```


```{r ggplot theme set}
theme_set(
  theme_minimal(base_size = 14)
)
```


```{r Utility function}
# Path of utils
utils.path <- "~/DPAS Reproducibility/utils"

# Calculate ICC on the total DPAS score
source(file.path(utils.path, "calculate_total_icc.R"))

# Calculate the ICC on each metric (lesion)
source(file.path(utils.path, "calculate_metric_icc.R"))
```


```{r data import}
# Path of raw data
path <- "~/DPAS Reproducibility/data"

# Read the DPAS48, 48 patients DPAS by 4 raters
raw_dpas <- read_excel(
  file.path(path, "DPAS48.xlsx")
)

# Read patient age and sex
raw_pt <- read_excel(
  file.path(path, "DPAS48_pt_info.xls")
)
```


```{r clean DPAS}
# Rename DPAS columns
clean_dpas <- raw_dpas |>
  rename(
    Patient_id = 患者,
    Rater = 监察人
  ) |>
  rename_with(
    # `rename_with` accepts function: 
    # dplyr/purrr accept one-sided function, function has only the RHS
    # `~...` R shorthand for an anonymous function 
    # `~` means: Define a function, and use `.x` as its input
    
    # `.x` represents the current input value passed to the function
    # Equivalent to: 
    # function(.x) {
    #   paste0("Forehead", sub("^额", "", .x))
    # }
    
    # Inside the paste0: 
    # for the current input column name .x
    # find 额 at the beginning of the string (^额)
    # and replace it with nothing ("")
    # leaving the rest of the column name unchanged.
    ~ paste0("Forehead", sub("^额", "", .x)),
    
    # select column start with `额`
    # followed by one or more digits
    matches("^额\\d+$")
  ) |>
  rename_with(
    ~ paste0("Right_cheek", sub("^右", "", .x)),
    matches("^右\\d+$")
  ) |>
  rename_with(
    ~ paste0("Left_cheek", sub("^左", "", .x)),
    matches("^左\\d+$")
  ) |>
  rename_with(
    ~ paste0("Chin", sub("^下", "", .x)),
    matches("^下\\d+$")
  )

# Long format
clean_dpas_long <- clean_dpas |>
  pivot_longer(
    cols = 3:ncol(clean_dpas),
    names_to = "Position_metric",
    values_to = "Score"
  ) |>
  separate(
    Position_metric,
    into = c("Position", "Metric"),
    sep = "(?<=\\D)(?=\\d)",
    convert = TRUE
  )

# Cleaning patient info
clean_pt <- raw_pt |>
  rename(
    Patient_id = 姓名,
    Sex = 性别,
    Age = 年龄
  ) |>
  mutate(
    Sex = recode(
      Sex,
      "男" = "Male",
      "女" = "Female"
    )
  )
```


```{r Merge DPAS and pt_info}
# Create canonical dataset, dat_long
# and join patient info
dat_long <- clean_pt |>
  left_join(
    clean_dpas_long,
    by = "Patient_id"
  )

# data class of each column 
dat_long <- dat_long %>%
  mutate(
    Patient_id = as.character(Patient_id),
    Sex        = factor(Sex),
    Age        = as.numeric(Age), # already numeric → no-op, kept for clarity
    Rater      = factor(Rater),
    Position   = factor(Position),
    Metric     = factor(Metric),
    Score      = as.numeric(Score) # already numeric → no-op, kept for clarity
  )
```


```{r revalue}
# Revalue metrics

# Revalue sex
dat_long <- dat_long |>
  mutate(Sex = case_when(
    Sex == "Female" ~ "女",
    Sex == "Male" ~ "男"
  ))

# 11 lesions metrics
dat_long <- dat_long |>
  mutate(Metric = case_when(
    Metric == "1" ~ "皮肤暗黄", 
    Metric == "2" ~ "黄色丘疹",
    Metric == "3" ~ "皮肤萎缩",
    Metric == "4" ~ "痣",
    Metric == "5" ~ "色素减退-色素沉着斑",
    Metric == "6" ~ "毛细血管扩张",
    Metric == "7" ~ "脂溢性角化",
    Metric == "8" ~ "老年性粉刺",
    Metric == "9" ~ "深皱纹",
    Metric == "10" ~ "浅表皱纹",
    Metric == "11" ~ "网状皱纹"
  ))

# Revalue the 4 positions
dat_long <- dat_long |>
  mutate(Position = case_when(
    Position == "Chin" ~ "下颌",
    Position == "Forehead" ~ "前额",
    Position == "Left_cheek" ~ "左颧骨",
    Position == "Right_cheek" ~ "右颧骨"
  ))
```

## 目的

    在皮肤科临床实践中需要稳定、量化的评估患者的诸多光老化皮损表现[@zhaoNoninvasiveAssessmentMethods2025]。本研究通过皮肤镜光老化量表（Dermoscopic Photoaging Scale, DPAS）[@isikDevelopmentSkinAging2013a]，探索在高原、高海拔区域的高老化皮损特征和性别差异。同时，通过组内相关系数分析（Intraclass Correlation Coefficient, ICC）[@IntraclassCorrelationsUses]探讨皮肤科不同职级人员实施DPAS量表的的可靠性和重现性。

## 方法

    随机选取皮肤科48名门诊患者，采集前额，左右颧骨区域和下巴正中的皮肤镜照片，按照DPAS量表要求，评估患者四个部位（前额、左右颧骨、下巴）的光老化相关皮损。科室四个职级人员（副主任医师、主治医师、住院医师、护士）分别按照DPAS量表要求读片。数据处理包括：

1.  统计不同光老化皮损在男女人群的分布特征
2.  采用组内相关系数（Intraclass Correlation Coefficient, ICC）[@kooGuidelineSelectingReporting31] 评估不同年资人员使用DPAS量表评估光老化的可靠性和重现性。
3.  采用线性混合效应逻辑回归（Linear mixed-effects logistic regression）探索性别、部位和不同光老化皮损的关系。

## 结果

### 1. 患者性别年龄


```{r Table 1}
table1_df <- dat_long |>
  dplyr::select(Patient_id, Sex, Age) |>
  dplyr::distinct(Patient_id, .keep_all = TRUE) |>
  gtsummary::tbl_summary(
    by = Sex,
    include = Age,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    missing = "ifany"
  ) |>
  gtsummary::add_p() |>
  tibble::as_tibble()

table1_df |>
  flextable() |>
  align(align = "center", part = "all") |>
  hline(
    i = nrow(table1_df),
    border = fp_border(width = 2),
    part = "body"
  ) |>
  set_caption("患者性别年龄") |>
  autofit()
```


### 2. 光老化皮损的发生率


```{r metric by sex,fig.width=7, fig.asp=.618, out.width="100%"}
score_sex_position_metric <- dat_long |>
  group_by(Sex, Position, Metric) |>
  summarise(n = sum(Score)/4) |>
  mutate(pct = n /48)


g <- ggplot(score_sex_position_metric,
            aes(x = fct_rev(Metric), y = pct, fill = Sex)) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = percent(pct, accuracy = 1)),
    position = position_dodge(width = 0.9),
    hjust = -0.1,          # push labels just past the bar tip
    size = 3
  ) +
  scale_fill_discrete_qualitative(palette = "Harmonic") + 
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.1)), # room for labels at the end
    limits = c(0, 0.5)
  ) +
  labs(
    title = "不同光老化皮损类型在男女四个部位的发生率",
    x = "光老化类型",
    y = "发生率",
    color = "性别"
  ) + 
  coord_flip() +
  facet_grid(. ~ Position) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "top",
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_blank()
  )

g
```


    如上图，毛细血管扩张、色斑，脂溢性角化和深浅皱纹是本研究人群的光老化特征，男女存在不同的皮损类型。


```{r total score}
dat_total_rater <- dat_long |>
  group_by(Patient_id, Rater, Sex, Age) |>
  summarise(
    Total = sum(Score, na.rm = TRUE),
    .groups = "drop"
  )
```


### 3. 患者DPAS总分的ICC重现性


```{r total icc}
icc_total <- calculate_total_icc(dat_total_rater)

icc_result <- as.data.frame(icc_total$results)

icc_result <- icc_result|>
  filter(type == "ICC2")%>% 
  select(ICC, p, `lower bound`, `upper bound`)

icc_result$p <- "<0.001"

icc_result |>
  flextable() |>
  align(align = "center", part = "all") |>
  hline(
    i = nrow(icc_result),
    border = fp_border(width = 2),
    part = "body"
  ) |>
  set_caption("四个评估人对患者DPAS总分的ICC可靠性") |>
  autofit()
```


### 4.患者分项皮损的ICC重现性


```{r metric specific score}
dat_metric_specific_rater <- dat_long |>
  group_by(Patient_id, Rater, Sex, Age, Metric) |>
  summarise(
    Total = sum(Score, na.rm = TRUE),
    .groups = "drop"
  )
```


```{r metric icc}
icc_metric <- calculate_metric_icc(dat_metric_specific_rater)

icc_metric$P <- "P < 0.001"

# Remove `皮肤萎缩`
icc_metric <- icc_metric |>
  filter(!Metric %in% "皮肤萎缩")

icc_metric |> 
  flextable() |>
  align(align = "center", part = "all") |>
  border_remove() |>
  hline(
    i = 1,
    border = officer::fp_border(width = 1),
    part = "header"
  ) |>
  hline_bottom(
    border = officer::fp_border(width = 2),
    part = "all"
  ) |>
  set_caption("四个评估人对患者分项皮损的ICC可靠性") |>
  autofit()
```

### 5. 混合效应逻辑回归

```{r dat_model}
dat_model <- dat_long |>
  mutate(
    Age_c = Age - mean(Age, na.rm = TRUE),
    Metric = factor(Metric),
    Position = factor(Position),
    Sex = factor(Sex)
  )

# Remove metric 3, which does not occur among all patients
# Inclusion of metric 3 lead to singular fitting
dat_model <- dat_model |>
  filter(!Metric == 3)
```


```{r glmer fit}
fit <- glmer(
  Score ~ Position + Age + Metric * Sex  + 
    (1 | Rater) + (1 | Patient_id),
  data = dat_model,
  family = binomial
)

# 3. Check convergence
# isSingular(fit)
# summary(fit)
```


```{r}
# emm_sex <- emmeans(
#   fit, 
#   ~ Sex | Metric, 
#   type = "response"
# )
# 
# pairs(emm_sex, reverse = TRUE)
```


```{r fig.width=6, fig.asp=.618, out.width="100%" plot prob}
# Estimated probabilities
emm_prob <- emmeans(
  fit,
  ~ Sex | Metric,
  type = "response"
) |>
  as.data.frame()

# Sex comparison within each metric
sex_p <- emmeans(
  fit,
  ~ Sex | Metric
) |>
  pairs() |>
  summary(infer = TRUE) |>
  as.data.frame() |>
  dplyr::mutate(
    p_label = dplyr::case_when(
      p.value < 0.001 ~ "p < 0.001",
      TRUE ~ paste0("p = ", format.pval(p.value, digits = 2, eps = 0.001))
    )
  ) |>
  dplyr::select(Metric, p_label)

# Order metrics by mean probability
metric_order <- emm_prob |>
  dplyr::group_by(Metric) |>
  dplyr::summarise(
    mean_prob = mean(prob),
    .groups = "drop"
  ) |>
  dplyr::arrange(mean_prob) |>
  dplyr::pull(Metric)

emm_prob <- emm_prob |>
  dplyr::mutate(
    Metric = factor(Metric, levels = metric_order)
  )

# Make p-value data frame use the same factor order
sex_p <- sex_p |>
  dplyr::mutate(
    Metric = factor(Metric, levels = metric_order)
  )

ggplot(
  emm_prob,
  aes(x = prob, y = Metric, color = Sex)
) +
  geom_errorbar(
    aes(
      xmin = asymp.LCL,
      xmax = asymp.UCL
    ),
    orientation = "y",
    position = position_dodge(width = 0.5),
    height = 0.15,
    linewidth = 0.4
  ) +
  geom_point(
    aes(shape = Sex),
    position = position_dodge(width = 0.5),
    size = 3
  ) +
  geom_hline(
    yintercept = seq(
      1.5,
      nlevels(emm_prob$Metric) - 0.5,
      by = 1
    ),
    color = "grey90",
    linewidth = 0.3,
    linetype = "dashed"
  ) +
  geom_text(
    data = sex_p,
    aes(
      x = 0.78,
      y = Metric,
      label = p_label
    ),
    inherit.aes = FALSE,
    hjust = 0,
    color = "black",
    size = 3.2
  ) +
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 0.90),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_color_discrete_qualitative(palette = "Harmonic") +
  labs(
    x = "发生率",
    y = "DPAS光老化类型",
    color = "性别",
    shape = "性别"
  ) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "top",
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

    逻辑回归结果表明，不同光老化皮损在男女人群的发生率不同，女性显著具有毛细血管扩张和色素色斑风险。男性光老化风险则是脂溢性角化和深浅皱纹。

## 结论

1. 在高原高海拔地区，毛细血管扩张、脂溢性角化和色素减退和分布不均是主要的光老化皮损。

2. 光老化皮损具有性别差异。

3. DPAS量表在临床门诊场景，具有很好稳定性和重现性。


## 参考文献
