---
title: "DPAS Reproducibility Thesis"
author: "Lian Rui"
date: "2026-09-24"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packgages

```{r}
# packages
library(tidyverse)
library(readxl)
library(dplyr)
library(tidyr)
library(psych) # ICC
library(ggplot2)
library(colorspace)
library(showtext)
showtext_auto()
library(GWalkR) # Tabuleu like, TODO: furrther explore
library(DataExplorer)
library(gtsummary) # Table 1
library(kableExtra)
```


## Background

## Methods

## Getting data

```{r data import}
path <- "~/DPAS Reproducibility/data"

# Import the DPAS48.xlsx
DPAS <- read_excel(paste0(path, "/", "DPAS48.xlsx"))

pt_info <- read_excel(paste0(path, "/", "DPAS48_pt_info.xls"))
```

## Cleaning data

```{r clean DPAS}
# The column names are in Chinese
# Translate to English as below: 
# 患者：Patient_id
# 监察人：Rater1, Rater12, Rater13, Rater14
# 前额: Forehead
# 右颊: Right cheek
# 左颊：Left cheek
# 下颌：Chin

DPAS <- DPAS |>
  rename("Patient_id" = "患者",
         "Rater" = "监察人")

# Use dplyr to change column names by pattern match in batch
# e.g., 额1 to 额11 as Forehead1 to forehead 11
DPAS <- DPAS |>
  rename_with(
    ~paste0("Forehead", sub("^额", "", .x)), 
    matches("^额\\d+$")
  )

DPAS <- DPAS |>
  rename_with(
    ~paste0("Right_cheek", sub("^右", "", .x)), 
    matches("^右\\d+$")
  )

DPAS <- DPAS |>
  rename_with(
    ~paste0("Left_cheek", sub("^左", "", .x)), 
    matches("^左\\d+$")
  )

DPAS <- DPAS |>
  rename_with(
    ~paste0("Chin", sub("^下", "", .x)), 
    matches("^下\\d+$")
  )
## TODO: a function to replace the four repeated action above!

# DPAS to long format
# Patient_id and Rater unchanged
DPAS <- DPAS |>
  pivot_longer(cols = 3:ncol(DPAS), 
               names_to = "Position_metric", 
               values_to = "Score")

# Separate column `Position_metric` into two columns
# one column of position and another position of metrics 1 to 11
DPAS <- DPAS |>
  separate(
    Position_metric, 
    into = c("Position", "Metric"), 
    sep = "(?<=\\D)(?=\\d)", 
    convert = TRUE
  )

```


```{r Cleaning pt_info}
# Change columns into English
# 姓名: Patient_id
# 性别: Sex
# 年龄：Age
pt_info <- pt_info |>
  rename(
    "Patient_id" = "姓名",
    "Sex" = "性别",
    "Age" = "年龄"
  )

# change Sex value: 
# 男: Male
# 女: Female
pt_info <- pt_info |>
  mutate(
    Sex = case_when(
      Sex == "男" ~ "Male", 
      Sex == "女" ~ "Female" 
    )
  )
```


```{r Merging}
# Merge table pt_info with table DPAS
dat <- pt_info |>
  left_join(DPAS, by = join_by("Patient_id"))
```

## TODO: other patient demographic, co-morbidity info maybe included

## Exploratory Analysis

### Table 1

```{r Table 1}
# Exposure var
library(gtsummary)
library(kableExtra)

table1_df <- dat |>
  dplyr::select(Patient_id, Sex, Age) |>
  dplyr::distinct(Patient_id, .keep_all = TRUE) |>
  gtsummary::tbl_summary(
    by = Sex,
    include = Age,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    missing = "ifany"
  ) |>
  gtsummary::add_p() |>
  tibble::as_tibble()

table1_df
```


### GWalkR
some exploratory analysis of raw data , any interactive package in R to explore dara with shiny

```{r GWalKR app }
gwalkr(data = dat)
```

### Distributions of score

```{r}
total_score <- dat |>
  group_by(Patient_id, Sex, Age, Rater) |>
  summarise(total = sum(Score, na.rm = TRUE)) |>
  ungroup()

# Visualize the distribution of total score
#| fig-width: 5
#| fig-height: 3
#| out-width: 70%
g<- ggplot(total_score, aes(x = total)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "grey") +
  labs(title = "Distribution of Total Score by rater and sex", x = "Total Score", y = "Count") +
  theme_minimal() + 
  facet_grid(Rater ~ Sex)

g
```


### Mean score by age, sex

```{r}
avg_score <- total_score |>
  group_by(Patient_id, Age, Sex) |>
  summarise(avg = mean(total, na.rm = TRUE))

avg_score

g <- ggplot(avg_score, aes(x = Age, y= avg, colour = Sex)) + 
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Mean DPAS Score by Age and Sex", x = "Age", y = "Mean DPAS Score") +
  theme_minimal()

g
```

## ICC evaluation of total score

```{r}
# convert the total_score data frame to wide format for ICC calculation
icc_data <- total_score %>%
  pivot_wider(
    # The names of the new columns
    names_from = Rater,
    
    # The values of the new columns
    values_from = total,
    
    # The prefix for the new columns
    names_prefix = "Rater"
  )

icc_data <- icc_data |>
  select(-Patient_id, -Age, -Sex)

# Calculate the icc using the psych package
icc_result <- ICC(icc_data) # Exclude the Patient column 

icc_result$results
```

## 对11项指标，三个评分人的评分是否存在主观差异

```{r}
# Calculate the ICC for each metric
# A data frame to store the ICC results
icc_results <- data.frame()

# Loop through each unique metric
for (m in seq_along(unique(dat$Metric))) {
  
  # tmp is a temporary data frame that contains the scores for the current metric
  tmp <- dat %>%
    # metric filtered as m in the loop
    filter(Metric == m) %>%
    group_by(Patient_id, Rater) %>%
    summarise(
      # for each metric, sum the scores across the 4 facial sites
      Score = sum(Score),    # sum across the 4 facial sites
      # remove the grouping structure after summarising
      .groups = "drop"
    ) %>%
    # Transform to wide format for ICC calculation
    pivot_wider(
      names_from = Rater,
      values_from = Score
    )

  icc <- ICC(tmp[, -1])
  # Store the ICC results in the icc_results data frame
  icc_results <- bind_rows(
    icc_results,
    data.frame(
      Metric = m,
      # Extract the ICC value for single random raters from the ICC results
      ICC2_1 = icc$results["Single_random_raters", "ICC"],
      
      # Extract the upper and lower bound of the ICC value 
      # for single random raters from the ICC results
      Lower = icc$results["Single_random_raters", "lower bound"],
      Upper = icc$results["Single_random_raters", "upper bound"],
      
      # p value
      p = icc$results["Single_random_raters", "p"]
    )
  )
}

icc_results
```


## Mixed effect model

```{r}
library(lmerTest)

fit <- lmer(
  total ~ Age + Sex + (1 | Rater) + (1 | Patient_id),
  data = total_score
)

```

```{r}
summary(fit)

```

```{r}
confint(fit, method = "Wald")
```


```{r}
VarCorr(fit)
```

