---
title: "DPAS Reproducibility Thesis"
author: "Lian Rui"
date: "2026-09-24"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Setup

### Packages

```{r}
# packages
library(tidyverse)
library(readxl)
library(dplyr)
library(tidyr)
library(psych) # ICC
library(ggplot2)
library(colorspace)
library(showtext)
showtext_auto()
library(GWalkR) # Tabuleu like, TODO: furrther explore
library(DataExplorer) 
library(gtsummary) # Table 1
library(kableExtra)
```

### Global options

### Helper functions


## 2. Data import

### Path

```{r path}
path <- "~/DPAS Reproducibility/data"
```

### raw_dpas

```{r raw_dpas import}
raw_dpas <- read_excel(
  file.path(path, "DPAS48.xlsx")
)
```

### raw_pt

```{r raw_pt import}
raw_pt <- read_excel(
  file.path(path, "DPAS48_pt_info.xls")
)
```

## 3. Data cleaning

### clean_dpas

```{r clean DPAS col names}
clean_dpas <- raw_dpas |>
  rename(
    Patient_id = 患者,
    Rater = 监察人
  ) |>
  rename_with(
    ~ paste0("Forehead", sub("^额", "", .x)),
    matches("^额\\d+$")
  ) |>
  rename_with(
    ~ paste0("Right_cheek", sub("^右", "", .x)),
    matches("^右\\d+$")
  ) |>
  rename_with(
    ~ paste0("Left_cheek", sub("^左", "", .x)),
    matches("^左\\d+$")
  ) |>
  rename_with(
    ~ paste0("Chin", sub("^下", "", .x)),
    matches("^下\\d+$")
  )
```

### clean_dpas_long

```{r clean DPAS long format}
clean_dpas_long <- clean_dpas |>
  pivot_longer(
    cols = 3:ncol(clean_dpas),
    names_to = "Position_metric",
    values_to = "Score"
  ) |>
  separate(
    Position_metric,
    into = c("Position", "Metric"),
    sep = "(?<=\\D)(?=\\d)",
    convert = TRUE
  )
```

### clean_pt

```{r rename revalue raw_pt}
clean_pt <- raw_pt |>
  rename(
    Patient_id = 姓名,
    Sex = 性别,
    Age = 年龄
  ) |>
  mutate(
    Sex = recode(
      Sex,
      "男" = "Male",
      "女" = "Female"
    )
  )
```

### Create canonical(single source of truth) dataset dat_long

```{r Merge DPAS and pt_info}
dat_long <- clean_pt |>
  left_join(
    clean_dpas_long,
    by = "Patient_id"
  )
```

## 4. Data quallity check

### Missing values

```{r}
sapply(dat_long, function(x) sum(is.na(x)))
```
### internal structure of dat_long

```{r dat_long structure}
str(dat_long)
```

### Change data type of each column

```{r change column class}
dat_long <- dat_long %>%
  mutate(
    Patient_id = as.character(Patient_id),
    Sex        = factor(Sex),
    Age        = as.numeric(Age), # already numeric → no-op, kept for clarity
    Rater      = factor(Rater),
    Position   = factor(Position),
    Metric     = factor(Metric),
    Score      = as.numeric(Score) # already numeric → no-op, kept for clarity
  )
```


### Data summary

```{r}
summary(dat_long)
```

## 5. Descriptive analysis

In `dat_long`, the columns are: 

```{r}
colnames(dat_long)
```
### Overall framework of descriptive analysis

We can define `Sex`, `Age`, `Rater`, `Position`, `Metric` as predictors and `Score` as outcome. and the description analysis in nature is combination of predictors and the distribution, mean, count etc of outcome. And such predictor combination has some hierarchy, e.g., single variable, bivariates and so on. Create a framework of description analysis

A general, combinatary structure as below: 

```text
Level 0: Outcome alone
        │
        ├── Overall distribution of Score
        │
Level 1: Single predictor
        │
        ├── Sex
        ├── Age
        ├── Rater
        ├── Position
        └── Metric
        │
Level 2: Two-dimensional descriptions
        │
        ├── Predictor × Score
        ├── Predictor × Predictor
        └── clinically/design-relevant combinations
        │
Level 3: Three-dimensional descriptions
        │
        ├── Sex × Position × Score
        ├── Sex × Metric × Score
        ├── Rater × Position × Score
        └── Rater × Metric × Score
        │
Level 4: Full / high-dimensional structure
        │
        └── Sex × Age × Rater × Position × Metric × Score
```

But this hierarchy makes the description analysis cumbersome and many of them are not scientific valid. 

### The scientific assumptions and questions behind the data

1. Sex related question: What are the difference of skin photoaging between male and female? 

1.1 Total score difference between male and female

1.2 The total score by facial area (chin, forehead, right-left cheek)

1.3 The overall lesion type difference

1.4 The lesion type by facial area

2. Age related question

2.1 The correlation of score with age and sex

2.2 The score of each facial area correlated with age and sex

2.3 the possible lesion type difference among young and old group and sex

3. Rater related question

3.1 The four scores of each patient by four raters

3.2 The overall distribution of total score by 4 raters

3.2 The distribution of 



### Table 1

```{r Table 1}
table1_df <- dat_long |>
  dplyr::select(Patient_id, Sex, Age) |>
  dplyr::distinct(Patient_id, .keep_all = TRUE) |>
  gtsummary::tbl_summary(
    by = Sex,
    include = Age,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    missing = "ifany"
  ) |>
  gtsummary::add_p() |>
  tibble::as_tibble()

table1_df
```


### Distributions of score by rater sex

```{r}
total_score_by_rater_sex <- dat |>
  group_by(Patient_id, Sex, Age, Rater) |>
  summarise(total = sum(Score, na.rm = TRUE)) |>
  ungroup()

# Visualize the distribution of total score
#| fig-width: 5
#| fig-height: 3
#| out-width: 70%
g<- ggplot(total_score_by_rater_sex, aes(x = total)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "grey") +
  labs(title = "Distribution of Total Score by rater and sex", x = "Total Score", y = "Count") +
  theme_minimal() + 
  facet_grid(Rater ~ Sex)

g
```


### Mean score by age, sex

```{r average score}
avg_score <- total_score_by_rater |>
  group_by(Patient_id, Age, Sex) |>
  summarise(avg = mean(total, na.rm = TRUE))
```

```{r avg score boxplot by sex}
g <- ggplot(avg_score, aes(x = Sex, y= avg, color = Sex)) + 
  geom_boxplot(show.legend = FALSE) + 
  geom_point(show.legend = FALSE) + 
  labs(title = "患者DPAS平均分的分布情况", 
       subtitle = "患者分数由四个评估人的分数平均得到",
       x = "性别",
       y = "DPAS 平均分") + 
  theme_minimal()

g
```

```{r compare mean by sex}
dpas_mean <- avg_score |>
  group_by(Sex) |>
  summarise(sex_mean = mean(avg, na.rm = TRUE), 
            sd = sd(avg, na.rm = TRUE))

g <- ggplot(dpas_mean, aes(x = Sex, y = sex_mean)) + 
  geom_col(aes(fill = Sex),
           show.legend = FALSE) + 
  geom_errorbar(aes(ymin = sex_mean - sd, ymax = sex_mean + sd), 
                width = 0.1) + 
  labs(title = "患者DPAS评分均值±SD",
       y = "患者DPAS评分均值",
       x = "性别") + 
  theme_minimal()

g
```


```{r correlation with age}

g <- ggplot(avg_score, aes(x = Age, y= avg, colour = Sex)) + 
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Mean DPAS Score by Age and Sex", x = "Age", y = "Mean DPAS Score") +
  theme_minimal()

g
```

## 男女在11项评分中的表现和差别

```{r}
avg_score_sex_metric <- dat |>
  group_by(Sex, Position, Metric) |>
  summarise(avg_score = mean(Score)) |>
  ungroup()


g <- ggplot(avg_score_sex_metric, 
            aes(x = Metric, y = avg_score, fill = Sex)) + 
  geom_col(position = position_dodge()) + 
  facet_wrap(Position~., nrow = 2) + 
  theme_minimal()

g
```

## ICC evaluation of total score

```{r}
# convert the total_score data frame to wide format for ICC calculation
icc_data <- total_score_by_rater_sex %>%
  pivot_wider(
    # The names of the new columns
    names_from = Rater,
    
    # The values of the new columns
    values_from = total,
    
    # The prefix for the new columns
    names_prefix = "Rater"
  )

icc_data <- icc_data |>
  select(-Patient_id, -Age, -Sex)

# Calculate the icc using the psych package
icc_result <- ICC(icc_data) # Exclude the Patient column 

icc_result$results
```

## 对11项指标，三个评分人的评分是否存在主观差异

```{r}
# Calculate ICC for each metric
# Each metric is first summed across the 4 facial regions
# for each patient and rater, producing a score from 0 to 4.

icc_results <- data.frame()

# Loop through each unique metric
for (m in unique(dat$Metric)) {
  
  # Create a temporary dataset for the current metric
  tmp <- dat %>%
    filter(Metric == m) %>%
    
    # For each patient and rater, sum the scores
    # across the 4 facial regions
    group_by(Patient_id, Rater) %>%
    summarise(
      Score = sum(Score),
      .groups = "drop"
    ) %>%
    
    # Convert raters from rows to columns
    pivot_wider(
      names_from = Rater,
      values_from = Score
    )

  # Calculate ICC
  icc <- ICC(tmp[, -1])
  
  # Extract ICC(2,1)
  icc_results <- bind_rows(
    icc_results,
    data.frame(
      Metric = m,
      ICC2_1 = icc$results[
        "Single_random_raters", 
        "ICC"
      ],
      Lower = icc$results[
        "Single_random_raters", 
        "lower bound"
      ],
      Upper = icc$results[
        "Single_random_raters", 
        "upper bound"
      ],
      p = icc$results[
        "Single_random_raters", 
        "p"
      ]
    )
  )
}

icc_results
```


## Mixed effect model

```{r}
library(lmerTest)

# 1. Center age
dat$Age_c <- as.numeric(scale(dat$Age, center = TRUE, scale = FALSE))

# convert metric as factor
dat$Metric <- factor(dat$Metric)

# 2. Fit model
fit <- glmer(
  Score ~ Metric +  Position +  Sex + Age_c + Metric * Sex + 
    (1 | Patient_id),
  data = dat,
  family = binomial
)

# 3. Check convergence
isSingular(fit)
summary(fit)
```

```{r}
print(x, correlation = TRUE)
```


```{r}
fit_main <- glmer(
  Score ~ Metric + Position + Sex + Age_c +
    (1 | Patient_id),
  data = dat,
  family = binomial
)

fit_interaction <- glmer(
  Score ~ Metric + Position + Sex + Age_c + Metric * Sex +
    (1 | Patient_id),
  data = dat,
  family = binomial
)

anova(fit_main, fit_interaction, test = "Chisq")
```

## Appendix: GWalkR fast exploration

Handy exploration of data using {GWalkR} package. 

```{r GWalKR app, viewer = "pane"}
# TODO: Specify the output to view pane, not inline? 
gwalkr(data = dat_long)
```