---
title: "DPAS Reproducibility Thesis"
author: "Lian Rui"
date: "2026-09-24"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r}
# packages
library(tidyverse)
library(readxl)
library(dplyr)
library(tidyr)
library(psych) # ICC
library(ggplot2)
library(colorspace)
library(showtext)
showtext_auto()
library(GWalkR) # Tabuleu like, TODO: furrther explore
library(DataExplorer) 
library(gtsummary) # Table 1
library(kableExtra)
```

## Background

## Methods

## Getting data

```{r data import}
path <- "~/DPAS Reproducibility/data"

# Import the DPAS48.xlsx
DPAS <- read_excel(paste0(path, "/", "DPAS48.xlsx"))

# Import pt info
pt_info <- read_excel(paste0(path, "/", "DPAS48_pt_info.xls"))
```

## Cleaning data

```{r clean DPAS col names}
# The column names are in Chinese
# Translate to English as below: 
# 患者：Patient_id
# 监察人：Rater1, Rater12, Rater13, Rater14
# 前额: Forehead
# 右颊: Right cheek
# 左颊：Left cheek
# 下颌：Chin
DPAS <- DPAS |>
  rename("Patient_id" = "患者",
         "Rater" = "监察人")

# Use dplyr to change column names by pattern match in batch
# e.g., 额1 to 额11 as Forehead1 to forehead 11
DPAS <- DPAS |>
  rename_with(
    ~paste0("Forehead", sub("^额", "", .x)), 
    matches("^额\\d+$")
  )

DPAS <- DPAS |>
  rename_with(
    ~paste0("Right_cheek", sub("^右", "", .x)), 
    matches("^右\\d+$")
  )

DPAS <- DPAS |>
  rename_with(
    ~paste0("Left_cheek", sub("^左", "", .x)), 
    matches("^左\\d+$")
  )

DPAS <- DPAS |>
  rename_with(
    ~paste0("Chin", sub("^下", "", .x)), 
    matches("^下\\d+$")
  )
## TODO: a function to replace the four repeated action above!
```


```{r clean DPAS long format}
# DPAS to long format
# Patient_id and Rater unchanged
DPAS <- DPAS |>
  pivot_longer(cols = 3:ncol(DPAS), 
               names_to = "Position_metric", 
               values_to = "Score")
```


```{r clean DPAS position_metric separation}
# Separate column `Position_metric` into two columns
# one column of position and another position of metrics 1 to 11
DPAS <- DPAS |>
  separate(
    Position_metric, 
    into = c("Position", "Metric"), 
    sep = "(?<=\\D)(?=\\d)", 
    convert = TRUE
  )
```


```{r Clean pt_info col names}
# Change columns into English
# 姓名: Patient_id
# 性别: Sex
# 年龄：Age
pt_info <- pt_info |>
  rename(
    "Patient_id" = "姓名",
    "Sex" = "性别",
    "Age" = "年龄"
  )
```


```{r clean pt_info sex value to English}
# change Sex value: 
# 男: Male
# 女: Female
pt_info <- pt_info |>
  mutate(
    Sex = case_when(
      Sex == "男" ~ "Male", 
      Sex == "女" ~ "Female" 
    )
  )
```


```{r Merge DPAS and pt_info}
# Merge table pt_info with table DPAS
dat <- pt_info |>
  left_join(DPAS, by = join_by("Patient_id"))
```

## TODO: other patient demographic, co-morbidity info maybe included

## Exploratory Analysis

### Table 1

```{r Table 1}
# Exposure var
library(gtsummary)
library(kableExtra)

table1_df <- dat |>
  dplyr::select(Patient_id, Sex, Age) |>
  dplyr::distinct(Patient_id, .keep_all = TRUE) |>
  gtsummary::tbl_summary(
    by = Sex,
    include = Age,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    missing = "ifany"
  ) |>
  gtsummary::add_p() |>
  tibble::as_tibble()

table1_df
```


### GWalkR
some exploratory analysis of raw data , any interactive package in R to explore dara with shiny

```{r GWalKR app, viewer = "pane"}
# TODO: Specify the output to view pane, not inline? 
gwalkr(data = dat)
```

### Distributions of score by rater sex

```{r}
total_score_by_rater_sex <- dat |>
  group_by(Patient_id, Sex, Age, Rater) |>
  summarise(total = sum(Score, na.rm = TRUE)) |>
  ungroup()

# Visualize the distribution of total score
#| fig-width: 5
#| fig-height: 3
#| out-width: 70%
g<- ggplot(total_score_by_rater_sex, aes(x = total)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "grey") +
  labs(title = "Distribution of Total Score by rater and sex", x = "Total Score", y = "Count") +
  theme_minimal() + 
  facet_grid(Rater ~ Sex)

g
```


### Mean score by age, sex

```{r average score}
avg_score <- total_score_by_rater |>
  group_by(Patient_id, Age, Sex) |>
  summarise(avg = mean(total, na.rm = TRUE))
```

```{r avg score boxplot by sex}
g <- ggplot(avg_score, aes(x = Sex, y= avg, color = Sex)) + 
  geom_boxplot(show.legend = FALSE) + 
  geom_point(show.legend = FALSE) + 
  labs(title = "患者DPAS平均分的分布情况", 
       subtitle = "患者分数由四个评估人的分数平均得到",
       x = "性别",
       y = "DPAS 平均分") + 
  theme_minimal()

g
```

```{r compare mean by sex}
dpas_mean <- avg_score |>
  group_by(Sex) |>
  summarise(sex_mean = mean(avg, na.rm = TRUE), 
            sd = sd(avg, na.rm = TRUE))

g <- ggplot(dpas_mean, aes(x = Sex, y = sex_mean)) + 
  geom_col(aes(fill = Sex),
           show.legend = FALSE) + 
  geom_errorbar(aes(ymin = sex_mean - sd, ymax = sex_mean + sd), 
                width = 0.1) + 
  labs(title = "患者DPAS评分均值±SD",
       y = "患者DPAS评分均值",
       x = "性别") + 
  theme_minimal()

g
```


```{r correlation with age}

g <- ggplot(avg_score, aes(x = Age, y= avg, colour = Sex)) + 
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Mean DPAS Score by Age and Sex", x = "Age", y = "Mean DPAS Score") +
  theme_minimal()

g
```

## 男女在11项评分中的表现和差别

```{r}
avg_score_sex_metric <- dat |>
  group_by(Sex, Position, Metric) |>
  summarise(avg_score = mean(Score)) |>
  ungroup()


g <- ggplot(avg_score_sex_metric, 
            aes(x = Metric, y = avg_score, fill = Sex)) + 
  geom_col(position = position_dodge()) + 
  facet_wrap(Position~., nrow = 2) + 
  theme_minimal()

g
```

## ICC evaluation of total score

```{r}
# convert the total_score data frame to wide format for ICC calculation
icc_data <- total_score_by_rater_sex %>%
  pivot_wider(
    # The names of the new columns
    names_from = Rater,
    
    # The values of the new columns
    values_from = total,
    
    # The prefix for the new columns
    names_prefix = "Rater"
  )

icc_data <- icc_data |>
  select(-Patient_id, -Age, -Sex)

# Calculate the icc using the psych package
icc_result <- ICC(icc_data) # Exclude the Patient column 

icc_result$results
```

## 对11项指标，三个评分人的评分是否存在主观差异

```{r}
# Calculate ICC for each metric
# Each metric is first summed across the 4 facial regions
# for each patient and rater, producing a score from 0 to 4.

icc_results <- data.frame()

# Loop through each unique metric
for (m in unique(dat$Metric)) {
  
  # Create a temporary dataset for the current metric
  tmp <- dat %>%
    filter(Metric == m) %>%
    
    # For each patient and rater, sum the scores
    # across the 4 facial regions
    group_by(Patient_id, Rater) %>%
    summarise(
      Score = sum(Score),
      .groups = "drop"
    ) %>%
    
    # Convert raters from rows to columns
    pivot_wider(
      names_from = Rater,
      values_from = Score
    )

  # Calculate ICC
  icc <- ICC(tmp[, -1])
  
  # Extract ICC(2,1)
  icc_results <- bind_rows(
    icc_results,
    data.frame(
      Metric = m,
      ICC2_1 = icc$results[
        "Single_random_raters", 
        "ICC"
      ],
      Lower = icc$results[
        "Single_random_raters", 
        "lower bound"
      ],
      Upper = icc$results[
        "Single_random_raters", 
        "upper bound"
      ],
      p = icc$results[
        "Single_random_raters", 
        "p"
      ]
    )
  )
}

icc_results
```


## Mixed effect model

```{r}
library(lmerTest)

# 1. Center age
dat$Age_c <- as.numeric(scale(dat$Age, center = TRUE, scale = FALSE))

# convert metric as factor
dat$Metric <- factor(dat$Metric)

# 2. Fit model
fit <- glmer(
  Score ~ Metric +  Position +  Sex + Age_c + Metric * Sex + 
    (1 | Patient_id),
  data = dat,
  family = binomial
)

# 3. Check convergence
isSingular(fit)
summary(fit)
```

```{r}
print(x, correlation = TRUE)
```


```{r}
fit_main <- glmer(
  Score ~ Metric + Position + Sex + Age_c +
    (1 | Patient_id),
  data = dat,
  family = binomial
)

fit_interaction <- glmer(
  Score ~ Metric + Position + Sex + Age_c + Metric * Sex +
    (1 | Patient_id),
  data = dat,
  family = binomial
)

anova(fit_main, fit_interaction, test = "Chisq")
```

