---
title: "Comorbidity in R"
author: "Lian Rui"
date: "2026-09-27"
output:
  pdf_document: default
  html_document: default
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Evidence shows skin disease and skin aging is associated with various co-morbidity.[@liLargescaleEpidemiologicalAnalysis2024]

Two approaches to include co-morbidity into the study:

1.  A composite comorbidity index: Charlson Comorbidity Index (CCI)[@charlsonNewMethodClassifying1987] and Elixhauser index[@elixhauserComorbidityMeasuresUse1998].
2.  Some common chronic diseases, CKD, Diabetes and risk factors (smoking)

## Objective

Establish the R skill to calculate some popular co-morbidity index

## Package

```{r}
# install package {comorbidity}
library(comorbidity)
library(tidyverse)
```

## Simulating ICD-10 code

It is also possible to simulate from two different versions of the ICD-10 coding system. The default is to simulate ICD-10 codes from the 2011 version:

```{r}
# load the comorbidity package
library(comorbidity)
# set a seed for reproducibility
set.seed(1)
# simulate 50 ICD-10 codes for 5 individuals
x <- data.frame(
  id = sample(1:5, size = 50, replace = TRUE),
  code = sample_diag(n = 50)
)

x <- x[order(x$id, x$code), ]

print(head(x, n = 15), row.names = FALSE)

```

## Compute co-morbidity score

```{r}
set.seed(1)
x <- data.frame(
  id = sample(1:3, size = 30, replace = TRUE),
  code = sample_diag(n = 30)
)
```

### Charlson

```{r}
charlson <- comorbidity(x = x, id = "id", code = "code", map = "charlson_icd10_quan", assign0 = TRUE) 
charlson
```

```{r}
cci <- score(charlson, weights = "charlson", assign0 = TRUE)
cci
```

### Elixhauser

```{r}
elixhauser <- comorbidity(x = x, id = "id", code = "code", map = "elixhauser_icd10_quan", assign0 = TRUE)
elixhauser
```

```{r}
exlixhauser_socore <- score(elixhauser, weights = "swiss", assign0 = TRUE)
exlixhauser_socore
```

### Questions

1.  Should the comorbidity index be summed up

Yes, in Charlson Index, each co-morbidity has weights and it is possible to assign levels of severity, e.g., malignancy. for exlixhauser, no weights but still has severity levels.

## Todo:

Verify the Version of ICD used

ICD10: 2009, 2011, 2017, 2018, 2022

ICD9: 2015?

## Summary

Beyond the well-documented effects of ultraviolet exposure, environmental factors such as air pollution and heat, lifestyle factors including smoking, sleep, and diet, and pathophysiological factors such as menopause, psychological stress, diabetes, and chronic kidney disease may all contribute to skin quality.[@addorPhotoagingAdditionalFactors2018a]

## Reference: 
