---
title: "Comorbidity in R"
author: "Lian Rui"
date: "2026-09-27"
output:
  pdf_document: default
  html_document: default
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Evidence shows skin disease and skin aging is associated with various co-morbidities[@liLargescaleEpidemiologicalAnalysis2024]

Two approaches to include co-morbiditties into the study:

1.  A composite comorbidity index: Charlson Comorbidity Index (CCI)[@charlsonNewMethodClassifying1987] and Elixhauser index[@elixhauserComorbidityMeasuresUse1998].
2.  Some common chronic diseases, CKD, Diabetes and risk factors (smoking)

## Objective

Establish the R skill to calculate some popular comorbidity index

## Package

```{r}
# install package comorbidity
library(comorbidity)
```

## Simulating ICD-10 code

It is also possible to simulate from two different versions of the ICD-10 coding system. The default is to simulate ICD-10 codes from the 2011 version:

```{r}
# load the comorbidity package
library(comorbidity)
# set a seed for reproducibility
set.seed(1)
# simulate 50 ICD-10 codes for 5 individuals
x <- data.frame(
  id = sample(1:5, size = 50, replace = TRUE),
  code = sample_diag(n = 50)
)

x <- x[order(x$id, x$code), ]

print(head(x, n = 15), row.names = FALSE)

```

## Compute comorbidity score

```{r}
set.seed(1)
x <- data.frame(
  id = sample(1:3, size = 30, replace = TRUE),
  code = sample_diag(n = 30)
)
```

### Charlson

```{r}
charlson <- comorbidity(x = x, id = "id", code = "code", map = "charlson_icd10_quan", assign0 = FALSE) 
charlson 
```

### Elixhauser

```{r}
elixhauser <- comorbidity(x = x, id = "id", code = "code", map = "elixhauser_icd10_quan", assign0 = FALSE)
elixhauser
```

## Todo:

Verify the Version of ICD used

ICD10: 2009, 2011, 2017, 2018, 2022

ICD9: 2015?

## Reference: 
