---
title: "Simulation -- ICC and Linear Mixed Model are Same Thing"
author: "Lian Rui"
date: "2026-07-31"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

A numeric simulation example in R to show the ICC results fro pysch package can be reproduced using linear mixed model and interpret the results from the two approach. 

## Setup

- 30 patients
- 3 raters
- patient SD = 2
- residual SD = 1
- no systematic rater bias

The model is 

$$
Y_{ij} = \mu + P_i + e_{ij}
$$

where 

$$
P_i \sim N(0, \sigma_P^2)
$$ 
and 

$$
e_{ij} \sim N(0, \sigma_e^2)
$$  

## Generate data

```{r}
library(tidyverse)

set.seed(123)

n_patient <- 30
n_rater   <- 3

sigma_patient <- 2
sigma_error   <- 1

patient_effect <-
  rnorm(n_patient,0,sigma_patient)

dat <-
  expand.grid(
    Patient=factor(1:n_patient),
    Rater=factor(1:n_rater)
  )

dat$Score <-
  10 +
  patient_effect[dat$Patient] +
  rnorm(nrow(dat),0,sigma_error)

head(dat)
```

## Compute ICC using psych package

```{r}
library(tidyr)
library(psych)

wide <-
  dat |>
  pivot_wider(
    names_from=Rater,
    values_from=Score
  )

ICC(wide[,-1])
```

### psych package reports many ICCS

| ICC   | Meaning             |
| ----- | ------------------- |
| ICC1  | One-way random      |
| ICC2  | Two-way random      |
| ICC3  | Two-way mixed       |
| ICC2k | Average of k raters |

#### Formula and summary of the six

| ICC      | ANOVA model    | Model equation                                                        | `lmer()` formula                            | Interpretation                                                                                                        |
| -------- | -------------- | --------------------------------------------------------------------- | ------------------------------------------- | --------------------------------------------------------------------------------------------------------------------- |
| ICC(1,1) | One-way random | (Score_{ij}=\mu+Patient_i+\epsilon_{ij})                              | `Score ~ 1 + (1 \| Patient)`                | Reliability of a single rating; raters are not modeled separately and rater variability is included in residual error |
| ICC(1,k) | One-way random | (Score_{ij}=\mu+Patient_i+\epsilon_{ij})                              | `Score ~ 1 + (1 \| Patient)`                | Reliability of the average score from k raters                                                                        |
| ICC(2,1) | Two-way random | (Score_{ij}=\mu+Patient_i+Rater_j+(Patient:Rater)*{ij}+\epsilon*{ij}) | `Score ~ 1 + (1 \| Patient) + (1 \| Rater)` | Single-rater absolute agreement; differences among raters are considered measurement error                            |
| ICC(2,k) | Two-way random | (Score_{ij}=\mu+Patient_i+Rater_j+(Patient:Rater)*{ij}+\epsilon*{ij}) | `Score ~ 1 + (1 \| Patient) + (1 \| Rater)` | Average of k raters; absolute agreement                                                                               |
| ICC(3,1) | Two-way mixed  | (Score_{ij}=\mu+Patient_i+Rater_j+\epsilon_{ij})                      | `Score ~ Rater + (1 \| Patient)`            | Single-rater consistency; systematic differences among these specific raters are ignored                              |
| ICC(3,k) | Two-way mixed  | (Score_{ij}=\mu+Patient_i+Rater_j+\epsilon_{ij})                      | `Score ~ Rater + (1 \| Patient)`            | Average of k raters; consistency among these specific raters                                                          |



## Fit the same data using LMM

```{r}
library(lme4)
# Fit the same data using LMM
library(lme4)

fit <-
  lmer(
    Score ~ 1 +
      (1|Patient),
    data=dat
  )

summary(fit)

VarCorr(fit)
```

### Compute ICC manually

```{r}
vc <- as.data.frame(VarCorr(fit))

sigma_patient <- vc$vcov[1]
sigma_error   <- vc$vcov[2]

ICC_lmm <-
  sigma_patient/
  (sigma_patient+sigma_error)

ICC_lmm
```
## Summary Connect the psych package and LMM

| ANOVA (psych)           | Linear Mixed Model                    |
| ----------------------- | ------------------------------------- |
| Patient mean square     | Patient random-effect variance        |
| Residual mean square    | Residual variance                     |
| ICC formula             | Variance partition coefficient        |
| F test                  | Likelihood/Wald test                  |
| Balanced data preferred | Handles unbalanced data naturally     |
| Produces ICC directly   | ICC computed from estimated variances |

## Shiny simulation

```{r}
library(shiny)
library(ggplot2)
library(lme4)
library(psych)
library(dplyr)
library(tidyr)

#---------------------------------------------------------
# UI
#---------------------------------------------------------

ui <- fluidPage(

  titlePanel("ICC and Linear Mixed Model Simulator"),

  sidebarLayout(

    sidebarPanel(

      sliderInput("sd_patient",
                  "Patient SD",
                  min = 0,
                  max = 5,
                  value = 2,
                  step = 0.1),

      sliderInput("sd_rater",
                  "Rater SD",
                  min = 0,
                  max = 3,
                  value = 0.5,
                  step = 0.1),

      sliderInput("sd_error",
                  "Residual SD",
                  min = 0.1,
                  max = 3,
                  value = 1,
                  step = 0.1),

      sliderInput("n_patient",
                  "Patients",
                  min = 10,
                  max = 100,
                  value = 30),

      sliderInput("n_rater",
                  "Raters",
                  min = 2,
                  max = 8,
                  value = 3),

      numericInput("seed",
                   "Random Seed",
                   value = 123),

      actionButton("simulate",
                   "Simulate")
    ),

    mainPanel(

      tabsetPanel(

        tabPanel("Observed Data",

                 plotOutput("scatter", height = 500)
        ),

        tabPanel("Variance",

                 plotOutput("variancePlot", height = 400),

                 tableOutput("varianceTable")
        ),

        tabPanel("ICC",

                 verbatimTextOutput("icc")
        ),

        tabPanel("Mixed Model",

                 verbatimTextOutput("lmm"))
      )
    )
  )
)

#---------------------------------------------------------
# SERVER
#---------------------------------------------------------

server <- function(input, output){

  dat <- eventReactive(input$simulate,{

    set.seed(input$seed)

    patient_effect <-
      rnorm(input$n_patient,
            0,
            input$sd_patient)

    rater_effect <-
      rnorm(input$n_rater,
            0,
            input$sd_rater)

    expand.grid(

      Patient=factor(1:input$n_patient),

      Rater=factor(1:input$n_rater)

    ) |>

      mutate(

        Score =
          10 +
          patient_effect[Patient] +
          rater_effect[Rater] +
          rnorm(n(),
                0,
                input$sd_error)
      )

  })

  fit <- reactive({

    lmer(

      Score ~
        (1|Patient)+
        (1|Rater),

      data=dat()

    )

  })

  #-------------------------------------------------------
  # Scatter
  #-------------------------------------------------------

  output$scatter <- renderPlot({

    ggplot(dat(),

           aes(Patient,
               Score,
               colour=Rater,
               group=Patient))+

      geom_line(alpha=.35)+

      geom_point(size=2)+

      theme_bw(base_size=15)+

      theme(axis.text.x=element_blank())

  })

  #-------------------------------------------------------
  # Variance
  #-------------------------------------------------------

  variance <- reactive({

    vc <-
      as.data.frame(VarCorr(fit()))

    data.frame(

      Component=c("Patient",
                  "Rater",
                  "Residual"),

      Variance=vc$vcov

    )

  })

  output$varianceTable <-

    renderTable({

      variance()

    },digits=3)

  output$variancePlot <- renderPlot({

    ggplot(variance(),

           aes(x="Variance",
               y=Variance,
               fill=Component))+

      geom_bar(stat="identity")+

      coord_flip()+

      theme_bw(base_size=16)

  })

  #-------------------------------------------------------
  # ICC
  #-------------------------------------------------------

  output$icc <- renderPrint({

    wide <-

      dat() |>

      pivot_wider(

        names_from=Rater,

        values_from=Score

      )

    ICC(wide[,-1])

  })

  #-------------------------------------------------------
  # Mixed Model
  #-------------------------------------------------------

  output$lmm <- renderPrint({

    summary(fit())

  })

}

shinyApp(ui,server)
```

